---
title: "SDS315HW_10"
output: html_document
date: "2024-04-21"
---
Name: Ankita Chitturi

EID: ac86269

Github:

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

redlining <- read.csv('redlining.csv')
groceries <- read.csv('groceries.csv')

library(ggplot2)
library(dplyr)
```
# Question 1

# Use a linear regression model to asses whether there is an association between the number of FAIR policies and the racial/ethnic composition of a ZIP code, adjusting for the fire, age, and income variables. 
```{r, echo=FALSE, include=FALSE}
red_reg <- lm(policies ~ minority + fire + age + income, data = redlining)
summary(red_reg)
```
Question: The question is whether there is an association between the number of fair policies and ethnic/racial composition of a ZIP code, adjusting for the age, fire, and income variable. 

Approach: I used multiple linear regression analysis and then included the number of fair policies as the dependent variable and the ethnic/racial composition, age variable, fire variable, and income variable as independent variables in the regression model. 

Results: When looking at at the association with racial/ethnic composition, the racial/ethnic composition of a ZIP code suggested a statistically significant relationship with the number of fair policies since the p-value is 0.0061. Moreover, for every one unit increase in ethnic/racial composition, there was an increase of 0.0083 fair policies. Additionally, when analyzing the association with fire incidents, the p-value is 0.0183 suggesting a statistically significant relationship. The estimand coefficient can conclude that for each unit increase in fire incidents, there was also an expected increase of around 0.021 fair policies. As for the age variable, the p-value is above 0.05 (it's 0.143), it can be concluded that there is no statistically significant association. Similarily, the income variable and its association is also not statistically significant as the p-value is approximately 0.678. This suggests that there is no significant relationship between income and age surrounding fair policies. 

Conclusion: These results indicate that factors which include but are not limited to ethnic/racial composition and fire incidents may affect the utilization of fair policies in a ZIP code. Stakehodlers who advocate for fairness and equity in insurance environments should at least consider these findings when creating interventions and policies. 

# Question 2

# Part A: What kind of price differences do we see across the different stores? Make a bar graph with Store on the vertical axis and average price of products sold at that store on the horizontal axis.
```{r, echo=FALSE}
avg_price <- groceries %>%
  group_by(Store) %>%
  summarise(average_price = mean(Price))

ggplot(avg_price, aes(x = Store, y = average_price)) +
  geom_bar(stat = "identity", color='black', fill='pink2') +
  coord_flip() +
  labs(title = "Average Price of Products Across Different Stores",
       x = "Store",
       y = "Average Price (in dollars)") + theme_classic()

```

The bar graph depicts the average prices of a myriad of products across various stores in multiple cities including Austin, Houston, and Fort Worth. Each single bar represents the average price of the products sold at that store, with the store names also listed. The average price is also shown on the x-axis. The graph itself allows the public to compare price patterns across a various types of stores and discuss variations and patterns in pricing of product strategies among these different retailers in these areas. 


# Part B: To illustrate this fact, please make a bar graph with Product on the vertical axis and number of stores selling that product on the horizontal axis.
```{r, echo=FALSE}
num_stores_per_product <- groceries %>%
  group_by(Product) %>%
  summarise(Num_Stores = n_distinct(Store))

ggplot(num_stores_per_product, aes(y = Product, x = Num_Stores)) +
  geom_bar(stat = "identity", color = 'black', fill = 'skyblue') +
  labs(title = "Number of Stores Selling Each Product",
       x = "Number of Stores",
       y = "Product") +
  theme_classic()
```

The bar graph shows the number of stores selling each product, depicting the variation in the availability of products across various retail outlets. Some products such as milk and eggs reach a maximum count of 16, which is a reflection of their widespread availability. On the other hand, products with lower counts suggests a narrower distribution among many stores.

# Part C: Now let’s use regression to try to isolate the effects of Type of store versus the actual products being sold. Fit a model for Price versus Product and the Type of store. 
```{r, echo=FALSE}
model <- lm(Price ~ Product + Type, data = groceries)

conf_int <- confint(model, level=0.95)
rounded_conf <- round(conf_int, 2)
 
head(rounded_conf, 10)
```
# Part D: Now fit a model for Price versus Product and Store. Which two stores seem to charge the lowest prices when comparing the same product? Which two stores seem to charge the highest prices when comparing the same product?
```{r, echo=FALSE}
model_store <- lm(Price ~ Product + Store, data = groceries)
coefs <- coef(model_store)
head(coefs, 10)

```
Kroger Fresh and Walmart seem to charge the lowest prices when comparing the same product. Conversely, Walgreens and Whole Foods are deemed to charge the highest prices when comparing the same product. 

# Part E: Let’s use your model from Part D to try to disambiguate between two possibilities:

# Possibility 1: central Market charges more than HEB for the same product.
# Possibiliy 2: central Market charges a similar amount to HEB for the same product.

# Try to put any difference between HEB and Central Market into the larger context: how big is the HEB/Central Market difference, compared to differences among other stores?

The coefficient for the Central Market is around -0.5734 and the coefficient for HEB is around -0.646. The central market charges slightly higher than HEB from looking merely at coefficients. However, there are more significant differences between other stores. For example, Fiesta has a coefficient of around -0.703 while the coefficient of Food Co-Op is approximately 0.290. This comparison suggests a more significant difference between Fiesta and Food Co-Op comparatively to HEB and Central Market. 

# Part F: Based on the sign of the Income10K coefficient, do consumers in poorer ZIP codes seem to pay more or less for the same product, on average? How large is the estimated size of the effect of Income10K on Price?
```{r, echo=FALSE}
groceries <- groceries %>%
  mutate(Income10K = Income / 10000)

model_income <- lm(Price ~ Product + Income10K, data = groceries)

coefs_income <- coef(model_income)

head(coefs_income, 10)

sd_income <- sd(groceries$Income10K)

sd_coefs <- coefs_income["Income10K"] * (sd(groceries$Price) / sd_income)
print(sd_coefs)

```

Since the coefficient for income is negative, this suggests that consumers in poor ZIP codes pay less for the same product. Through the standardized income coefficient, it can be concluded that one-standard deviation increase in the income of ZIP code seems to be associated with a -0.006 standard deviation change in the price that consumers in the ZIP code expect to pay for the same product. 


